Robust Spectroscopic Inference with Imperfect Models

### 2014-12-06 tex/paper.tex
Coauthor comments addressed.

@@

All astronomers recognize that spectroscopy offers a wealth of information that can help
characterize the properties of the observing target.  In the context of stellar astrophysics,
spectroscopy plays many fundamental roles.  The relative strengths and widths of stellar absorption
lines provide access to key parameters like effective temperature ($T_{\rm eff}$) and surface
gravity ($\log g$), enabling model comparisons in the Hertzsprung-Russell diagram to estimate the
masses and ages so crucial to understanding stellar evolution, as well as individual elemental (and
molecular) abundances or the collective ``metallicity" (typically parameterized as [Fe/H]),
facilitating study of the chemical hallmarks of different stellar populations.  With sufficient
spectral resolution, the velocity content of a spectrum can convey crucial information about stellar
rotation ($v \sin i$) and kinematics (e.g., association with a cluster or companion through the
radial velocity, $v_z$).  While many fields benefit from the spectroscopic measurements of these
stellar properties, they are of acute interest to the exoplanet community.  There, all estimates of
the planet properties are made {\it relative} to the host properties (e.g., the planet-to-host mass
function or radius {\it ratio} is constrained with the radial velocity or transit techniques,
respectively).  Moreover, essential clues to the planet formation process are encapsulated in the
dependences of planet frequency on host mass \citep[e.g.,][]{johnson07,howard10} and metallicity
\citep[e.g.,][]{fischer05,buchhave14}.

@@

We opt for the computationally expedient approach that employs a library of model spectra, 
$\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, 
[{\rm Fe/H}]]$.  However, it is worth noting that the techniques developed here are applicable to 
{\it any} ``back-end" that generates a model spectrum.  In our adopted approach, the model spectrum 
for an arbitrary $\vt_{\ast}$ is interpolated from a spectral library,
\begin{equation} \label{eqn:interp} 
\flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto \flam(\vt_{\ast}), 
\end{equation} 
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency
is critical.  In practice, a simple linear interpolation from nearby grid points in the spectral
library is suitably fast, but introduces a non-negligible level of interpolation error. Although
generally small, this error can significantly bias the inference of stellar parameters
when the signal-to-noise of the data is sufficiently high. To account for the uncertainty in a given
interpolated spectrum, we design a Bayesian spectral emulator, which we describe in detail in
Appendix \ref{sec:Appendix}. Rather than delivering a single interpolated spectrum for a given
$\vt_\ast$, the emulator delivers a probability distribution over a set of possible interpolated
spectra. By marginalizing over all of the possible interpolated models, we can modify the likelihood
function (introduced in Section~\ref{subsec:likelihood}) to account for the interpolation uncertainty. For
simplicity, however, we now proceed as if we used linear interpolation from a synthetic library to
generate model spectra and that the error introduced by this process is negligible, since the
following concepts are applicable regardless of the interpolation method.

@@

Generally, the ``raw" model spectrum $\flam(\vt_{\ast})$ is highly over-sampled compared to the 
observed spectrum, and does not account for several additional observational and instrumental 
effects that become important in comparisons with real data.  Therefore, a certain amount of 
post-processing is required before assessing the model quality.  We treat that post-processing in 
two stages: the first deals with an additional set of ``observational" parameters, $\vt_{\rm obs}$, 
that incorporate dynamical effects, geometry, and the relative location of the target, while the 
second employs a suite of nuisance parameters ($\vp$) designed to mitigate an imperfect 
data calibration.

@@

\begin{figure}[b!]
\includegraphics{figs/chebyshev.pdf}
\vspace{-0.75cm}
\figcaption{A demonstration of our treatment for residual calibration mismatches.  The observed 
spectra at the overlap of two echelle orders ({\it top}) have slightly ($\sim$1--3\%) discrepant 
shapes.  By using Chebyshev polynomials ({\it middle, top}) one can correct for that mismatch by
adjusting the data ({\it middle, bottom}); instead, in practice we equivalently distort the model
by these polynomials ({\it bottom}) such that the model remains linear in the Chebyshev coefficients
(Eq.~\ref{eqn:chebyshev}).  \label{fig:chebyshev}}
\end{figure}

@@

We employ an analogous, but more rigorous, approach to deal with this issue, cast {\it internally}
to the modeling framework to propagate the uncertainties introduced by additional degrees of freedom
in the model.  Deviations in the spectral shape introduced by calibration residuals are treated as
explicit contributions to the model spectrum, enabling an exploration of the distribution of
possible ``tweaks" to the calibration that can be compartmentalized into a set of nuisance
parameters and then marginalized.  This way, the inferences for the stellar parameters will properly
account for the underlying uncertainty in the calibration of the spectral shape.  In practice, this
is achieved by distorting segments of the model spectrum with polynomials, $\mathsf{P}$
\citep[e.g.,][]{eisenstein06,koleva09}. Figure \ref{fig:chebyshev} offers a practical demonstration
of how these nuisance parameters are applied to the model. For data that has $N_{\rm ord}$ spectral
orders, each denoted with index $o$, the model spectrum can be decomposed into the set of orders
\begin{eqnarray} \label{eqn:chebyshev}
\vM(\vT, \cheb) &=& \{ \vM_o(\vT) \times \mathsf{P}_o \} \\
                &=& \{ \vM_o(\vT) \times \sum_n c_o^{(n)} \, T_o^{(n)} \}, \nonumber
\end{eqnarray}
where $T^{(n)}$ is an $n^{\rm th}$ degree Chebyshev function.  The $n \times N_{\rm ord}$ 
coefficients are considered a set of nuisance parameters, $\cheb = [\{c_o^{(0)}, c_o^{(1)}, 
\ldots, c_o^{(n-1)} \}]$.  With judicious priors, we can ensure that the unintended treatment of 
real spectral features (e.g., molecular bands) as residual calibration artifacts is negligible.  
The lowest-degree (scaling) coefficient, $c^{(0)}$, is naturally degenerate with the solid angle, 
$\Omega$.  Therefore, we enforce an additional constraint that the mean of the polynomial is 
unity.  For data with a single spectral order, this means simply setting $c^{(0)} = 1$.  In the 
multiple order case, we assign $c^{(0)} = 1$ in an arbitrary order as an anchor, but permit the 
$c^{(0)}$ in other orders to be different.

@@

The quality of the model spectrum is assessed by comparing to the data with a pixel-by-pixel 
likelihood calculation.  If we denote the data spectrum as $\vD$, then a corresponding residual 
spectrum (an $N_{\rm pix}$-element vector) can be defined for any input parameter set,
\begin{equation}
\vR \equiv \vR(\vT, \cheb) \equiv \vD-\vM(\vT, \cheb).
\end{equation}
To quantify the probability of the data conditioned on the model, we adopt a standard 
multi-dimensional Gaussian likelihood function
\begin{equation}
p(\vD|\vM) =  \frac{1}{[(2 \pi)^{N_{\rm pix}} \det(\vC)]^{1/2}} \exp\left ( -\frac{1}{2} \,
   \vR^\trans \vC^{-1} \vR \right )
   \label{eqn:likelihood}
\end{equation}
that penalizes models which yield larger residuals and explicitly allows for covariances in the 
residual spectrum through the $N_{\rm pix} \times N_{\rm pix}$ matrix $\vC$.  For practical 
reasons, the log-likelihood is used as the quality metric, where
\begin{equation}
  \ln{p(\vD | \vM)} = -\frac{1}{2} \left( \vR^\trans \vC^{-1} \vR + \ln{\det{\vC}} + N_{\rm pix} \ln{2\pi} \right).
  \label{eqn:lnlikelihood}
\end{equation}

@@

Astronomical spectrographs are designed such that the detector over-samples the instrumental 
line-spread function with at least a few pixels.  Therefore, adjacent pixels never record 
independent samples of the true spectrum.  In that case, a difference between an observed and 
modeled spectral feature will create a correlated residual that spans multiple pixels.  This can be 
demonstrated practically through the autocorrelation of $\vR$: a slight model mismatch will produce 
correlated residuals over a characteristic scale similar to the instrumental broadening kernel 
width ($\sigma_v$).  Figure \ref{fig:class0} highlights a specific example of these correlated 
residuals in real data; a significant autocorrelation signal is found on a $\sim$5 pixel scale, 
corresponding to the 6.8\,km s$^{-1}$ FWHM of $\mathcal{F}_v^{\rm inst}$.

\begin{figure}[!htb]
\begin{center}
  \includegraphics{residuals_23.pdf}
  \includegraphics{class0_residuals.pdf}
  \includegraphics{class0_autocorrelation.pdf}
  \figcaption{({\it top}) A comparison of the data and a typical model with parameters drawn from 
the posterior distribution, along with the corresponding residual spectrum.  ({\it middle}) A 
zoomed view of the gray band in the top panels, highlighting the mildly covariant residual 
structure that is produced by slight mismatches between the data and model spectra.  ({\it bottom}) 
The autocorrelation of the residual spectrum.  Notice the substantial autocorrelation signal for 
offsets of $\lesssim 8$ pixels, demonstrating clearly that the residuals are not well described by 
white (Poisson) noise alone. \label{fig:class0}}
\end{center}
\end{figure}

@@

For a well-designed spectrograph and sufficiently accurate model, this {\it global} (i.e., present 
throughout the spectrum) covariance should have a relatively low amplitude and small 
correlation length.  To describe that structure, we use a stationary covariance kernel (or radial 
basis function) with an amplitude that depends only on the velocity distance between two pixels, 
\begin{equation}
  r_{ij} = r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
   - \lambda_j}{ \lambda_i + \lambda_j} \right |,
\end{equation}
where $c$ is the speed of light.  This parametric kernel describes the covariance between pixel 
residuals, 
\begin{equation}
  \Kglobal_{ij} =  \langle \vR_i \; \vR_j \rangle.
  \label{eqn:expectation}
\end{equation}
A variety of kernels have been used in the field of Gaussian processes to parameterize such a 
covariant structure \citep[e.g.,][]{rasmussen05}.  Here, we adopt the commonly-employed \matern\ 
kernel with $\nu = 3/2$,
\begin{equation}
  \Kglobal_{ij}(\vp_{{\mathsf C}, {\rm g}}) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
   \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
   \label{eqn:global}
\end{equation}
where the hyperparameters $\vp_{{\mathsf C}, {\rm g}} = [a_{\rm g}, \ell]$ include an amplitude 
($a_{\rm g}$) and scale ($\ell$).  To ensure that $\vC$ remains a relatively sparse matrix that 
enables computational expediency, we employ a Hann window function
\begin{equation}
  w_{ij}^{\rm G} \,(r_0) = \left \{ 
    \begin{array}{cc}
    \frac{1}{2} + \frac{1}{2} \cos \left(\frac{\pi r_{ij}}{r_0} \right) & r_{ij} \le r_0 \\
    0 & r_{ij} > r_0 \\
  \end{array}
  \right .
  \label{eqn:Hann}
\end{equation}
as a taper.  The truncation distance $r_0$ can be fixed to a reasonable multiple of the scale (we 
set $r_0 = 4\ell$).

@@

Assuming that these local residuals are produced primarily by pathological differences in the 
spectral line strength (rather than shape or center), a simple Gaussian is a reasonable residual 
model.  In that case, the $k^{\rm th}$ such local residual could be described as
\begin{equation}
\vR = a_k \exp \left[ - \frac{r^2(\lambda,\mu_k)}{2\sigma_k^2} \right],
\end{equation}
with peak amplitude $a_k$, mean wavelength $\mu_k$, and width $\sigma_k$.  Following 
Eq.~\ref{eqn:expectation}, the covariance of any two pixels related to the $k^{\rm th}$ local 
residual regions can be written
\begin{equation} \label{eqn:kregion}
  \mathcal{K}^{\textrm{L},k}_{ij} = a_k^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ],
\end{equation}
such that the full local covariance kernel composed of a sequence of residual regions is the linear 
combination,
\begin{equation} \label{eqn:klocal}
  \Klocal_{ij}(\vp_{{\mathsf C},l}) = \sum_k w^k_{ij} \, \mathcal{K}^{\textrm{L},k}_{ij},
\end{equation}
with a set of hyperparameters $\vp_{{\mathsf C},\textrm{L}} = [\{a_k, \mu_k, \sigma_k\}]$.  Note that we 
again taper the kernels with Hann windows (Eq.~\ref{eqn:Hann}) to ensure a sparse covariance 
matrix; in this case, the truncation distance $r_0$ can be set to some multiple of the width 
parameter (e.g., $r_0 = 4\sigma_k$).  In effect, these kernels systematically down-weight the 
influence of strong residuals in the likelihood calculation, mitigating any potential bias they 
might induce on inferences of the interesting parameters ($\vT$).  This acts like a robust, 
flexible, and unbiased method for (correlated) outlier rejection that preserves the integrity of 
the probabilistic framework (as opposed to the common manual or threshold-based techniques of 
clipping or masking).

@@

We can now compute the covariance matrix employed in the likelihood calculation 
(Eq.~\ref{eqn:lnlikelihood}) as the linear combination of the trivial pixel-by-pixel noise matrix 
and the global and local kernels discussed above, 
\begin{equation}
\vC_{ij}(\cov)  = b \, \delta_{ij} \, \sigma_i + w_{ij} \, \Kglobal_{ij}(\vp_{{\mathsf C}, g}) + 
                  \Klocal_{ij}(\vp_{{\mathsf C}, l}), 
\end{equation}
with hyperparameters $\cov = [\vp_{{\mathsf C}, \textrm{G}}, \vp_{{\mathsf C}, \textrm{L}}]$.  The factor $b$ is 
a parameter that scales up the Poisson noise in each pixel by a constant factor to account for 
additional detector or data reduction uncertainties (e.g., read noise, uncertainties in the 
spectral extraction procedure, etc.); typically $b < 1.1$ for well-calibrated optical spectra.  If 
there are $N_{\rm loc}$ local covariance patches (see Section \ref{subsec:MCMC} on how this is 
determined), then there are $4N_{\rm loc}+2$ elements in the set of covariance hyperparameters, 
$\cov$.

Figure \ref{fig:matrix} provides a graphical illustration of how the kernels that comprise the 
covariance matrix are able to reproduce the structure present in a typical residual spectrum (for 
the sake of clarity, the interpolation error kernel is discussed separately, in the Appendix).

@@

@@

We suggest assuming uniform priors on the global kernel parameters.  For the local kernels, we 
typically adopt flat priors for the amplitudes and means \{$a_k$, $\mu_k$\}, but construct a 
specialized prior for the widths \{$\sigma_k$\} that is flat below the instrumental broadening 
width (when $\sigma_k < \sigma_v$) and has a rapid (logistic) taper to zero at larger values.  This 
form is designed to prevent the local kernels from diffusing to large $\sigma_k$ and low $a_k$,
since such behavior is better treated by the global kernel.

\begin{figure}[!htb]
  \includegraphics{figs/PGM.pdf}
  \figcaption{ Probabilistic graphical model.
\label{fig:PGM}}
\end{figure}

@@

The inference problem developed here has a natural hierarchical structure between the collections 
of ``interesting" parameters, $\vT = [\vt_{\ast}, \vt_{\rm obs}]$, and ``nuisance" hyperparameters, 
$\vP = [\cheb, \cov]$.  To explore the posterior distribution of the model conditioned on the data
\begin{equation} 
p(\vT, \vP | \vD) \propto p(\vD | \vT, \vP) \, p(\vT, \vP)
\label{eqn:post}
\end{equation}
for this type of structure, it is convenient to use MCMC simulations with a blocked Gibbs 
sampler coupled to the Metropolis-Hastings algorithm.  To summarize, this procedure works by 
sampling in a subset of parameters (with Metropolis-Hastings) conditioned on the current (fixed) 
values of the other parameters; after each iteration, the Gibbs sampler then updates the values of 
the sampled parameter subset and then cycles through all the (previously fixed) different parameter 
subsets in the same way (for a more nuanced mathematical description of the process, see Chapter 11 
of \citealt{gelman13}).  A step-by-step prescription follows, where the $i^{\rm th}$ 
iteration of the Gibbs sampler is indexed with a superscript: \\

>>>>

\hyphenation{ExoLab}
\hyphenation{ExoStat}

@@

All astronomers recognize that spectroscopy offers a wealth of information that can help
characterize the properties of the observing target.  In the context of stellar astrophysics,
spectroscopy plays many fundamental roles.  The relative strengths and widths of stellar absorption
lines provide access to key parameters like effective temperature ($T_{\rm eff}$) and surface
gravity ($\log g$), enabling model comparisons in the Hertzsprung-Russell diagram to estimate the
masses and ages so crucial to understanding stellar evolution, as well as individual elemental (and
molecular) abundances or the collective ``metallicity" (typically parameterized as [Fe/H]),
facilitating study of the chemical hallmarks of different stellar populations.  With sufficient
spectral resolution, the velocity content of a spectrum can convey crucial information about stellar
rotation ($v \sin i$) and kinematics (e.g., association with a cluster or companion through the
radial velocity, $v_z$).  While many fields benefit from the spectroscopic measurements of these
stellar properties, they are of acute interest to the exoplanet community.  There, all estimates of
the planet properties are made {\it relative} to the host properties (e.g., the planet-to-host
radius {\it ratio} or mass function is constrained with the radial velocity or transit techniques,
respectively).  Moreover, essential clues to the planet formation process are encapsulated in the
dependences of planet frequency on host mass \citep[e.g.,][]{johnson07,howard10} and metallicity
\citep[e.g.,][]{fischer05,buchhave14}.

@@

We opt for the computationally expedient approach that employs a library of model spectra, 
$\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, 
[{\rm Fe/H}]]$.  However, it is worth noting that the techniques developed here are applicable to 
{\it any} ``back-end" that generates a model spectrum.  In our adopted approach, the model spectrum 
for an arbitrary $\vt_{\ast}$ is interpolated from a spectral library,
\begin{equation} \label{eqn:interp} 
\flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto \flam(\vt_{\ast}), 
\end{equation} 
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency
is critical.  In practice, a simple linear interpolation from nearby grid points in the spectral
library is suitably fast, but introduces a non-negligible level of interpolation error. Although
generally small, this error can significantly bias the inference of stellar parameters
when the signal-to-noise of the data is sufficiently high. To account for the uncertainty in a given
interpolated spectrum, we design a Bayesian spectral emulator, which we describe in detail in
Appendix \ref{sec:Appendix}. Rather than delivering a single interpolated spectrum for a given
$\vt_\ast$, the emulator delivers a probability distribution over a set of possible interpolated
spectra. By marginalizing over all of the possible interpolated models, we can modify the likelihood
function (introduced in Section~\ref{subsec:likelihood}) to account for the interpolation
uncertainty. However, since the following concepts are applicable regardless of the interpolation
method, for simplicity we now proceed as if we were using linear interpolation to generate model spectra
and assume that the error introduced by this process is negligible.

@@

Generally, the ``raw" model spectrum $\flam(\vt_{\ast})$ is highly over-sampled compared to the 
observed spectrum, and does not account for several additional observational and instrumental 
effects that become important in comparisons with real data.  Therefore, a certain amount of 
post-processing is required before assessing the model quality.  We treat that post-processing in 
two stages: the first deals with an additional set of ``observational" parameters, $\vt_{\rm obs}$, 
that incorporate dynamical effects, geometry, and the relative location of the target, while the 
second employs a suite of nuisance parameters $\vp$ designed to mitigate an imperfect 
data calibration.

@@

@@

\begin{figure}[t!]
\includegraphics{figs/chebyshev.pdf}
\vspace{-0.75cm}
\figcaption{A demonstration of our treatment for residual calibration mismatches.  The observed 
spectra at the overlap of two echelle orders ({\it top}) have slightly ($\sim$1--3\%) discrepant 
shapes.  By using Chebyshev polynomials ({\it middle, top}) one can correct for that mismatch by
adjusting the data ({\it middle, bottom}); instead, in practice we equivalently distort the model
by these polynomials ({\it bottom}) such that the model remains linear in the Chebyshev coefficients
(Eq.~\ref{eqn:chebyshev}).  \label{fig:chebyshev}}
\end{figure}

We employ an analogous, but more rigorous, approach to deal with this issue, cast {\it internally}
within the modeling framework to propagate the uncertainties introduced by additional degrees of freedom
in the model.  Deviations in the spectral shape introduced by calibration residuals are treated as
explicit contributions to the model spectrum, enabling an exploration of the distribution of
possible ``tweaks" to the calibration that can be compartmentalized into a set of nuisance
parameters and then marginalized.  This way, the inferences for the stellar parameters will properly
account for the underlying uncertainty in the calibration of the spectral shape.  In practice, this
is achieved by distorting segments of the model spectrum with polynomials, $\mathsf{P}$
\citep[e.g.,][]{eisenstein06,koleva09}. Figure \ref{fig:chebyshev} offers a practical demonstration
of how these nuisance parameters are applied to the model. For data that has $N_{\rm ord}$ spectral
orders, each denoted with index $o$, the model spectrum can be decomposed into the set of orders
\begin{eqnarray} \label{eqn:chebyshev}
\vM(\vT, \cheb) &=& \{ \vM_o(\vT) \times \mathsf{P}_o \} \\
                &=& \{ \vM_o(\vT) \times \sum_n c_o^{(n)} \, T_o^{(n)} \}, \nonumber
\end{eqnarray}
where $T^{(n)}$ is an $n^{\rm th}$ degree Chebyshev function.  The $n \times N_{\rm ord}$ 
coefficients are considered a set of nuisance parameters, $\cheb = [\{c_o^{(0)}, c_o^{(1)}, 
\ldots, c_o^{(n-1)} \}]$.  With judicious priors, we can ensure that the unintended treatment of 
real spectral features (e.g., molecular bands) as residual calibration artifacts is negligible.  
The lowest-degree (scaling) coefficient, $c^{(0)}$, is naturally degenerate with the solid angle, 
$\Omega$.  Therefore, we enforce an additional constraint that the mean of the polynomial is 
unity.  For data with a single spectral order, this means simply setting $c^{(0)} = 1$.  In the 
multiple order case, we assign $c^{(0)} = 1$ in an arbitrary order as an anchor, but permit the 
$c^{(0)}$ in other orders to be different.

\vspace{1cm}

@@

The quality of the model spectrum is assessed by comparing to the data with a pixel-by-pixel 
likelihood calculation.  If we denote the data spectrum as $\vD$, then a corresponding residual 
spectrum (an $N_{\rm pix}$-element vector) can be defined for any input parameter set,
\begin{equation}
\vR \equiv \vR(\vT, \cheb) \equiv \vD-\vM(\vT, \cheb).
\end{equation}
To quantify the probability of the data conditioned on the model, we adopt a standard 
multi-dimensional Gaussian likelihood function
\begin{equation}
p(\vD|\vM) =  \frac{1}{[(2 \pi)^{N_{\rm pix}} \det(\vC)]^{1/2}} \exp\left ( -\frac{1}{2} \,
   \vR^\trans \vC^{-1} \vR \right )
   \label{eqn:likelihood}
\end{equation}
that penalizes models which yield larger residuals and explicitly allows for covariances in the
residual spectrum through the $N_{\rm pix} \times N_{\rm pix}$ matrix $\vC$\footnote{When using a
spectral emulator to interpolate model spectra, this matrix will be further modified to account
for the propagated interpolation uncertainty (Appendix \ref{sec:Appendix}, Eq.~\ref{eqn:modC})}.
For practical reasons, the log-likelihood is used as the quality metric, where
\begin{equation}
  \ln{p(\vD | \vM)} = -\frac{1}{2} \left( \vR^\trans \vC^{-1} \vR + \ln{\det{\vC}} + N_{\rm pix} \ln{2\pi} \right).
  \label{eqn:lnlikelihood}
\end{equation}

@@

Astronomical spectrographs are designed such that the detector over-samples the instrumental 
line-spread function with at least a few pixels.  Therefore, adjacent pixels never record 
independent samples of the true spectrum.  In that case, a difference between an observed and 
modeled spectral feature will create a correlated residual that spans multiple pixels.  This can be 
demonstrated practically through the autocorrelation of $\vR$: a slight model mismatch will produce 
correlated residuals over a characteristic scale similar to the instrumental broadening kernel 
width ($\sigma_v$).  Figure \ref{fig:class0} highlights a specific example of these correlated 
residuals in real data; a significant autocorrelation signal is found on an $\sim$8 pixel scale, 
corresponding to the 6.8\,km s$^{-1}$ FWHM of $\mathcal{F}_v^{\rm inst}$.

\begin{figure}[!t]
\begin{center}
  \includegraphics{residuals_23.pdf}
  \includegraphics{class0_residuals.pdf}
  \includegraphics{class0_autocorrelation.pdf}
  \figcaption{({\it top}) A comparison of the data and a typical model with parameters drawn from 
the posterior distribution, along with the corresponding residual spectrum.  ({\it middle}) A 
zoomed view of the gray band in the top panels, highlighting the mildly covariant residual 
structure that is produced by slight mismatches between the data and model spectra.  ({\it bottom}) 
The autocorrelation of the residual spectrum.  Notice the substantial autocorrelation signal for 
offsets of $\lesssim 8$ pixels, demonstrating clearly that the residuals are not well described by 
white (Poisson) noise alone. \label{fig:class0}}
\end{center}
\end{figure}

@@

For a well-designed spectrograph and sufficiently accurate model, this {\it global} (i.e., present 
throughout the spectrum) covariance should have a relatively low amplitude and small 
correlation length.  To describe that structure, we use a stationary covariance kernel (or radial 
basis function) with an amplitude that depends only on the velocity distance between two pixels, 
\begin{equation}
  r_{ij} = r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
   - \lambda_j}{ \lambda_i + \lambda_j} \right |,
\end{equation}
where $c$ is the speed of light.  This parametric kernel describes the covariance between pixel 
residuals, 
\begin{equation}
  \Kglobal_{ij} =  \langle \vR_i \; \vR_j \rangle.
  \label{eqn:expectation}
\end{equation}
A variety of kernels have been used in the field of Gaussian processes to parameterize such a 
covariant structure \citep[e.g.,][]{rasmussen05}.  Here, we adopt the commonly-employed \matern\ 
kernel with $\nu = 3/2$,
\begin{equation}
  \Kglobal_{ij}(\vp_{{\mathsf C}, {\rm G}}) = a_{\rm G} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
   \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
   \label{eqn:global}
\end{equation}
where the hyperparameters $\vp_{{\mathsf C}, {\rm G}} = [a_{\rm G}, \ell]$ include an amplitude
($a_{\rm G}$) and scale ($\ell$). We distinguish the kernel parameters from other parameters in the
model by labelling them hyperparameters. Because a Gaussian process describes a population of
functions that can be generated by random draws from the covariance matrix, the covariance kernel
parameters are truly part of a hierarchical model. In our case, the functions described by these
hyperparameters are actually the many different realizations of covariant residuals that result from
a spectral fit. For an example of the Gaussian process kernel and the population of covariant
residual ``noise'' that could be generated from it, see Figure~\ref{fig:matrix}.
To ensure that $\vC$ remains a relatively sparse matrix that enables computational expediency, we
employ a Hann window function
\begin{equation}
  w_{ij} \,(r_0) = \left \{ 
    \begin{array}{cc}
    \frac{1}{2} + \frac{1}{2} \cos \left(\frac{\pi r_{ij}}{r_0} \right) & r_{ij} \le r_0 \\
    0 & r_{ij} > r_0 \\
  \end{array}
  \right .
  \label{eqn:Hann}
\end{equation}
to taper the kernel.  The truncation distance $r_0$ can be fixed to a reasonable multiple of the scale (we 
set $r_0 = 4\ell$).

@@

Assuming that these local residuals are produced primarily by pathological differences in the 
spectral line strength (rather than shape or center), a simple Gaussian is a reasonable residual 
model.  In that case, the $k^{\rm th}$ such local residual could be described as
\begin{equation}
\vR = a_k \exp \left[ - \frac{r^2(\lambda,\mu_k)}{2\sigma_k^2} \right],
\end{equation}
with peak amplitude $a_k$, mean wavelength $\mu_k$, and width $\sigma_k$.  Following 
Eq.~\ref{eqn:expectation}, the covariance of any two pixels related to the $k^{\rm th}$ local 
residual regions can be written
\begin{equation} \label{eqn:kregion}
  \mathcal{K}^{\textrm{L},k}_{ij} = a_k^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ],
\end{equation}
such that the full local covariance kernel composed of a sequence of residual regions is the linear 
combination,
\begin{equation} \label{eqn:klocal}
  \Klocal_{ij}(\vp_{{\mathsf C},L}) = \sum_k^N w^k_{ij} \, \mathcal{K}^{\textrm{L},k}_{ij},
\end{equation}
with a set of hyperparameters $\vp_{{\mathsf C},\textrm{L}} = [\{a_1, \mu_1, \sigma_1, \ldots, a_N,
\mu_N, \sigma_N\}]$.  Note that we again taper the kernels with Hann windows (Eq.~\ref{eqn:Hann}) to ensure
a sparse covariance matrix; in this case, the truncation distance $r_0$ can be set to some multiple
of the width parameter (e.g., $r_0 = 4\sigma_k$).  In effect, these kernels systematically
down-weight the influence of strong residuals in the likelihood calculation, mitigating any
potential bias they might induce on inferences of the interesting parameters ($\vT$).  This acts
like a robust, flexible, and unbiased method for (correlated) outlier rejection that preserves the
integrity of the probabilistic framework (as opposed to the common manual or threshold-based
techniques of masking or clipping).

Alternatively, one can also think of adding in this Gaussian component to the covariance matrix as
being equivalent to marginalizing over the existence of an outlier spectral line with a prior on the
line strength that is strictly Gaussian, where adjusting the hyperparameters of the local kernel
changes the prior on the strength and shape of the outlier residual. We find that once such a local
kernel is instantiated at the location of an outlier spectral line, the inference on the stellar
parameters ($\vT$) is moderately robust to the precise values of the kernel parameters. This means
that the marginalization over outlier lines can be performed at minimal computational cost, since
MCMC iterations which would otherwise be spent exploring the parameter space of the local kernel
hyperparameters are largely unnecessary if one is only interested in the stellar parameters. For
completeness, however, the local kernel hyperparameters of the spectral fits presented in
Section~\ref{sec:examples} are allowed to float and explored simultaneously with the other stellar
and nuisance parameters.

@@

We can now compute the covariance matrix employed in the likelihood calculation 
(Eq.~\ref{eqn:lnlikelihood}) as the linear combination of the trivial pixel-by-pixel noise matrix 
and the global and local kernels discussed above, 
\begin{equation}
\vC_{ij}(\cov)  = b \, \delta_{ij} \, \sigma_i + w_{ij} \, \Kglobal_{ij}(\vp_{{\mathsf C}, G}) + 
                  \Klocal_{ij}(\vp_{{\mathsf C}, L}), 
\end{equation}
with hyperparameters $\cov = [\vp_{{\mathsf C}, \textrm{G}}, \vp_{{\mathsf C}, \textrm{L}}]$.  The factor $b$ is 
a parameter that scales up the Poisson noise in each pixel by a constant factor to account for 
additional detector or data reduction uncertainties (e.g., read noise, uncertainties in the 
spectral extraction procedure, etc.); typically $b < 1.1$ for well-calibrated optical spectra.  If 
there are $N_{\rm loc}$ local covariance patches (see Section \ref{subsec:MCMC} on how this is 
determined), then there are $4N_{\rm loc}+2$ elements in the set of covariance hyperparameters, 
$\cov$.

Figure \ref{fig:matrix} provides a graphical illustration of how the kernels that comprise the 
covariance matrix are able to reproduce the structure present in a typical residual spectrum.

@@

\begin{figure}[!b]
  \includegraphics{figs/PGM.pdf}
  \figcaption{A probabilistic graphical model representing how the parameters of the model are
    combined to forward model a spectrum.  When interpolating models using a spectral emulator
    (Appendix \ref{sec:Appendix}), the stellar parameters ($\vt_\ast$) determine the weights
    ($\mathbf{w}$) of the eigenspectra ($\mathbf{\Xi}$), which are modified according to the
    observational parameters ($\vt_\textrm{obs}$) and polynomial parameters ($\cheb$). Together,
    these parameters specify the model spectrum ($\vM$). If one uses linear interpolation instead of
    a spectral emulator, then there would be no intermediate nodes for $\mathbf{w}$ and $\mathbf{\Xi}$. The structure
    of the covariance matrix, which is included in the likelihood function, is determined by the
    covariance hyperparameters ($\cov$). Together, the model spectrum and the covariance matrix
    predict the resulting dataset ($\vD$).
\label{fig:PGM}}
\end{figure}

@@

We suggest assuming uniform priors on the global kernel parameters.  For the local kernels, we
typically adopt flat priors for the amplitudes and means \{$a_k$, $\mu_k$\}, but construct a
specialized prior for the widths \{$\sigma_k$\} that is flat below the combined broadening width
(when $\sigma_k$ is less than the FWHM of $\varphi$) and has a rapid (logistic) taper to zero at
larger values.  This form is designed to prevent the local kernels from diffusing to large
$\sigma_k$ and low $a_k$, since such behavior is better treated by the global kernel.

@@

The inference problem developed here has a natural structure between the collections of
``interesting" parameters, $\vT = [\vt_{\ast}, \vt_{\rm obs}]$, the nuisance calibration parameters
$\cheb$, and the covariance hyperparameters $\cov$. The conditional dependencies of these parameters
are shown graphically in Figure~\ref{fig:PGM}. To explore the posterior distribution of the model
conditioned on the data
\begin{equation} 
p(\vT, \cheb, \cov | \vD) \propto p(\vD | \vT, \cheb, \cov) \, p(\vT, \cheb, \cov)
\label{eqn:post}
\end{equation}
for this type of structure, it is convenient to use MCMC simulations with a blocked Gibbs 
sampler coupled to the Metropolis-Hastings algorithm.  To summarize, this procedure works by 
sampling in a subset of parameters (with Metropolis-Hastings parameter proposals) conditioned on the current (fixed) 
values of the other parameters; after each iteration, the Gibbs sampler then updates the values of 
the sampled parameter subset and then cycles through all the (previously fixed) different parameter 
subsets in the same way (for a more nuanced mathematical description of the process, see Chapter 11 
of \citealt{gelman13}).  A step-by-step prescription follows, where the $i^{\rm th}$ 
iteration of the Gibbs sampler is indexed with a superscript: \\

\noindent (1) Initialize the parameters.  One might set $\vT^0$ based on estimates in the 
literature or scaling behaviors, and make simple assumptions about $\cheb^0$.  Here, we set the 
Chebyshev coefficients ($\cheb^0$) so that the polynomials are constant ($c_o^{(0)} = 1$,  
$c_o^{(>0)} = 0$, $\forall$ $o$) and assume only the trivial noise spectrum and interpolation error 
kernel contribute to the covariance matrix ($\cov^0 = 0$).  \\
